---
title: "Econ 43 Final"
Date: 12/5/17
output: html_document
---
#Names: Gianluca, Jeffrey, & Jonathan


```{r setup, include=FALSE}
#==============================================================================
#   Transparent California
#==============================================================================

# original by Michael Kevane 10/14/2017
# editied by Gianluca, Jeffrey, & Jonathan

# Description: Create tables of descriptive statistics for 
# Salary data for public employees in California
# data is from http://transparentcalifornia.com

#==============================================================================
#   1. Settings, packages, and options
#==============================================================================

# Clear the working space
rm(list = ls())

#install.packages("viridis", "raster", "ggmap", "mapproj", "maps", "maptools", "mapdata", "sp", "ggplot2", "dplyr")

#install.packages(c("maps", "mapdata"))
#install.packages("dplyr")
#install.packages("reshape")
#install.packages("ggrepel")
#install.packages("noncensus")
#install.packages("plotly")


# Load packages
library(plyr)
library(sp) 
library(raster)
library(viridis)
library(leaflet)
library(doBy)
library(dplyr)
library(foreign)
library(gdata)
library(ggplot2)
library(sandwich)
library(stargazer)
library(tidyr)
library(maps)
library(mapdata)
library(mapproj)
library(maptools)
library(ggmap)
library(reshape2)
library(reshape)
library(ggrepel)
library(noncensus)
library(knitr)
library(plotly)


# turn off scientific notation except for big numbers
options(scipen = 9)
# function to calculate corrected SEs for regression 
cse = function(reg) {
  rob = sqrt(diag(vcovHC(reg, type = "HC1")))
  return(rob)
}
```

```{r Data, include=FALSE}

#==============================================================================
#   2. Data section
#==============================================================================

### Read data 

# Data input using read.csv
# Note, albany data available for 2011-2016
# many other cities, counties, agencies available

# Data input using read.csv
sonoma_c <- read.csv("http://transparentcalifornia.com/export/sonoma-county-2016.csv")
napa_c <- read.csv("http://transparentcalifornia.com/export/napa-county-2016.csv")
solano_c <- read.csv("http://transparentcalifornia.com/export/solano-county-2016.csv")
marin_c <- read.csv("http://transparentcalifornia.com/export/marin-county-2016.csv")
contra_costa_c <- read.csv("http://transparentcalifornia.com/export/contra-costa-county-2016.csv")
san_fran_c <- read.csv("https://transparentcalifornia.com/export/san-francisco-2016.csv")
san_mateo_c <- read.csv("http://transparentcalifornia.com/export/san-mateo-county-2016.csv")
santa_cruz_c <- read.csv("http://transparentcalifornia.com/export/santa-cruz-county-2016.csv")
santa_clara_c <- read.csv("http://transparentcalifornia.com/export/santa-clara-county-2016.csv")
alameda_c <- read.csv("http://transparentcalifornia.com/export/alameda-county-2016.csv")

merged.data <- rbind(sonoma_c,napa_c,solano_c,marin_c,contra_costa_c,san_fran_c,san_mateo_c,santa_cruz_c,santa_clara_c,alameda_c, all=TRUE)

#Average Sonoma County base pay over $50k
sonoma_c$basepay = tolower(sonoma_c$Total.Pay...Benefits)
son50 <- sonoma_c$basepay[sonoma_c$Total.Pay...Benefits > 50000]
son50 <- as.numeric(son50)
son50 <- matrix(data = son50, ncol = 1)
son_avg <- mean(son50)
num_son <- nrow(son50)
tot_son <- nrow(sonoma_c)

#Average Napa County base pay over $50k
napa_c$basepay = tolower(napa_c$Total.Pay...Benefits)
napa50 <- napa_c$basepay[napa_c$Total.Pay...Benefits >50000]
napa50 <- as.numeric(napa50)
napa50 <- matrix(data = napa50, ncol=1)
napa_avg <- mean(napa50)
num_napa <- nrow(napa50)
tot_napa <- nrow(napa_c)

#Average Solano County base pay over $50k
solano_c$basepay = tolower(solano_c$Total.Pay...Benefits)
solano50 <- solano_c$basepay[solano_c$Total.Pay...Benefits >50000]
solano50 <- as.numeric(solano50)
solano50 <- matrix(data = solano50, ncol = 1)
solano_avg <- mean(solano50)
num_solano <- nrow(solano50)
tot_solano <- nrow(solano_c)

#Average Marin base pay over $50k
marin_c$basepay = tolower(marin_c$Total.Pay...Benefits)
marin50 <- marin_c$basepay[marin_c$Total.Pay...Benefits > 50000]
marin50 <- as.numeric(marin50)
marin50 <- matrix(data = marin50, ncol = 1)
marin_avg <- mean(marin50)
num_marin <- nrow(marin50)
tot_marin <- nrow(marin_c)

#Average Contra Costa County base pay over $50k
contra_costa_c$basepay = tolower(contra_costa_c$Total.Pay...Benefits)
cc50 <- contra_costa_c$basepay[contra_costa_c$Total.Pay...Benefits > 50000]
cc50 <- as.numeric(cc50)
cc50 <- matrix(data = cc50, ncol = 1)
cc_avg <- mean(cc50)
num_cc <- nrow(cc50)
tot_cc <- nrow(contra_costa_c)

#Average San Francisco County base pay over $50k
san_fran_c$basepay = tolower(san_fran_c$Total.Pay...Benefits)
sf50 <- san_fran_c$basepay[san_fran_c$Total.Pay...Benefits > 50000]
sf50 <- as.numeric(sf50)
sf50 <- matrix(data = sf50, ncol = 1)
sf_avg <- mean(sf50)
num_sf <- nrow(sf50)
tot_sf <- nrow(san_fran_c)

#Average San Mateo County base pay over $50k
san_mateo_c$basepay = tolower(san_mateo_c$Total.Pay...Benefits)
san_mateo50 <- san_mateo_c$basepay[san_mateo_c$Total.Pay...Benefits > 50000]
san_mateo50 <- as.numeric(san_mateo50)
san_mateo50 <- matrix(data = san_mateo50, ncol = 1)
san_mateo_avg <- mean(san_mateo50)
num_mateo <- nrow(san_mateo50)
tot_mateo <- nrow(san_mateo_c)

#Average Santa Cruz County base pay over $50k
santa_cruz_c$basepay = tolower(santa_cruz_c$Total.Pay...Benefits)
cruz50 <- santa_cruz_c$basepay[santa_cruz_c$Total.Pay...Benefits > 50000]
cruz50 <- as.numeric(cruz50)
cruz50 <- matrix(data = cruz50, ncol = 1)
cruz_avg <- mean(cruz50)
num_cruz <- nrow(cruz50)
tot_cruz <- nrow(santa_cruz_c)

#Average Santa Clara County base pay over $50k
santa_clara_c$basepay = tolower(santa_clara_c$Total.Pay...Benefits)
clara50 <- santa_clara_c$basepay[santa_clara_c$Total.Pay...Benefits >50000]
clara50 <- as.numeric(clara50)
clara50 <- matrix(data = clara50, ncol = 1)
clara_avg <- mean(clara50)
num_clara <- nrow(clara50)
tot_clara <- nrow(santa_clara_c)

#Average Alameda County base pay over $50k
alameda_c$basepay = tolower(alameda_c$Total.Pay...Benefits)
alameda50 <- alameda_c$basepay[alameda_c$Total.Pay...Benefits > 50000]
alameda50 <- as.numeric(alameda50)
alameda50 <- matrix(data = alameda50, ncol = 1)
alameda_avg <- mean(alameda50)
num_alam <- nrow(alameda50)
tot_alam <- nrow(alameda_c)

son_avg
napa_avg
solano_avg
marin_avg
cc_avg
sf_avg
san_mateo_avg
cruz_avg
clara_avg
alameda_avg

data(counties)
counties
sonoma_pop <-subset(counties, state == "CA" & county_name == "Sonoma County", select = c(county_name, population))
napa_pop <- subset(counties, state == "CA" & county_name == "Napa County", select = c(county_name, population))
solano_pop <-subset(counties, state == "CA" & county_name == "Solano County", select = c(county_name, population))
marin_pop <- subset(counties, state =="CA" & county_name == "Marin County", select = c(county_name, population))
cc_pop <- subset(counties, state =="CA" & county_name == "Contra Costa County", select = c(county_name, population))
sf_pop <- subset(counties, state == "CA" & county_name=="San Francisco County", select = c(county_name, population))
mateo_pop <- subset(counties, state == "CA" & county_name == "San Mateo County", select = c(county_name, population))
cruz_pop <- subset(counties, state =="CA" & county_name == "Santa Cruz County", select = c(county_name, population))
clara_pop <- subset(counties, state =="CA" & county_name == "Santa Clara County", select = c(county_name, population))
alam_pop <- subset(counties, state == "CA" & county_name == "Alameda County", select = c(county_name, population))

```

```{r Manipulates Data, include=FALSE,warning=FALSE, Messege=FALSE, echo=FALSE,comment=NA}

cities <- c("Sonoma","Napa", "Solano", "Marin", "Contra Costa", "San Francisco", "San Mateo", "Santa Cruz","Santa Clara","Alameda")
geocode(cities)

#Inserts values into separate Data Frames (each data frame is one column)
city_rats <- c(num_son/tot_son, num_napa/tot_napa, num_solano/tot_solano, num_marin/tot_marin, 
               num_cc/tot_cc, num_sf/tot_sf, num_mateo/tot_mateo, num_cruz/tot_cruz, 
               num_clara/tot_clara, num_alam/tot_alam)
city_rats <- as.data.frame(city_rats)
city_names <- as.data.frame(cities)
city_locs <- as.data.frame(geocode(cities))

#Merge data frames horizontally
all_together <- as.data.frame(c(city_names, city_rats, city_locs))

#Rename the columns
colnames(all_together) <- c("CityName", "AveragePay", "Longitude", "Latitude")
all_together <- all_together[, -c(5:7)]
all_together

```

```{r Maps, echo=FALSE, ,Message= FALSE, warning=FALSE, Messege=FALSE}

#Makes Maps
myLocation <- "San Francisco"

myMap <- get_map(location=myLocation,
                 source="google", maptype="roadmap", zoom=9)

RatMap <- ggmap(myMap)+ 
  geom_point(data = all_together, mapping = aes(x = Longitude, y = Latitude, colour = AveragePay), size = 7) + scale_colour_gradient(low = "#132B43", high = "#FF8C00")  +   labs(x = 'Longitude', y = 'Latitude') +
  geom_label_repel(data = all_together, aes(x = Longitude, y = Latitude, label = CityName), size=3,
                   fill = "white", box.padding = unit(.5, "lines"),
                   label.padding = unit(.2, "lines"),
                   segment.color = "red", segment.size = 1) 

RatMap +labs(colour = "County Percent")  + guides(size=FALSE) + 
  ggtitle("Percent of Public Employees with Total Pay above $50K to Total Number of Public Employees") +
  theme(plot.title = element_text(family = "Times", color="#666666", face="bold", size=9, hjust=0)) +
  theme(axis.title = element_text(family = "Times", color="#666666", face="bold", size=10)) 
```

## Reasoning/Analysis

For this data, we wanted to examine what counties in the bay area would be the best place for a new government employee to make a work in. To find this, we first subset the data to be all salaries over $50,000. We did this for two reasons: 1)To eliminate part-time jobs 2) To look for what we thought would be a reasonable minimum living wage in the bay area. 

Afer performing this subset, we examined the percentage of government employees in each of these counties making above that $50,000 mark. This was performed by dividing the number of employees making over #50,000 by total number government employees in the area. Then, as shown by the map, the best places to work happened to be Alameda, Solano and Santa Clara.  


```{r 2nd Graph, echo=FALSE, warning=FALSE, Messege=FALSE}
#Create graph to see how much Benefits/Base Pay contribute to Total Pay (plus benefits) in each county
ben_son <- mean(sonoma_c$Benefits)
base_son <-mean(sonoma_c$Base.Pay)
total_son <- mean(sonoma_c$Total.Pay...Benefits)

ben_napa <- mean(napa_c$Benefits)
base_napa <- mean(napa_c$Base.Pay)
total_napa <- mean(napa_c$Total.Pay...Benefits)

ben_sol <- mean(solano_c$Benefits)
base_sol <- mean(solano_c$Base.Pay)
total_sol <- mean(solano_c$Total.Pay...Benefits)

ben_marin <- mean(marin_c$Benefits)
base_marin <- mean(marin_c$Base.Pay)
total_marin <- mean(marin_c$Total.Pay...Benefits)

ben_cc <- mean(contra_costa_c$Benefits)
base_cc <- mean(contra_costa_c$Base.Pay)
total_cc <- mean(contra_costa_c$Total.Pay...Benefits)

ben_sf <- as.numeric(san_fran_c$Benefits)
ben_sf <- mean(ben_sf)
base_sf <- as.numeric(san_fran_c$Base.Pay)
base_sf <- mean(base_sf)
total_sf <- mean(san_fran_c$Total.Pay...Benefits)

ben_mateo <- mean(san_mateo_c$Benefits)
base_mateo <- mean(san_mateo_c$Base.Pay)
total_mateo <- mean(san_mateo_c$Total.Pay...Benefits)

ben_cruz <- mean(santa_cruz_c$Benefits)
base_cruz <- mean(santa_cruz_c$Base.Pay)
total_cruz <- mean(santa_cruz_c$Total.Pay...Benefits)

ben_clara <- mean(santa_clara_c$Benefits)
base_clara <-mean(santa_clara_c$Base.Pay)
total_clara <- mean(santa_clara_c$Total.Pay...Benefits)

ben_alam <- mean(alameda_c$Benefits)
base_alam <- mean(alameda_c$Base.Pay)
total_alam <- mean(alameda_c$Total.Pay...Benefits)

benefit_df <- as.data.frame(c(ben_son, ben_napa, ben_sol, ben_marin, ben_cc, ben_sf, ben_mateo, ben_cruz, ben_clara, ben_alam))
base_df <- as.data.frame(c(base_son, base_napa, base_sol, base_marin, base_cc, base_sf, base_mateo, base_cruz, base_clara, base_alam))
total_df <- as.data.frame(c(total_son, total_napa, total_sol, total_marin, total_cc, total_sf, total_mateo, total_cruz, total_clara, total_alam))
pays_tog <- as.data.frame(c(city_names, base_df, benefit_df, total_df))
colnames(pays_tog) <- c("County.Name", "Base.Pay", "Benefits", "Total.Pay")
pays_tog <- pays_tog[, -c(5:7)]

pays_tog <- pays_tog[order(pays_tog$Total.Pay),]

p <- plot_ly(pays_tog, x = ~Benefits, y = ~County.Name, name = "Benefits" , type = 'scatter',
             mode = "markers", marker = list(color = "green"),
             hoverinfo = 'text',
             text = ~paste('Avg. Benefits: $', Benefits)) %>%
  add_trace(x = ~Base.Pay, y = ~County.Name, name = "Base Pay", type = 'scatter',
            mode = "markers", marker = list(color = "blue"),
            hoverinfo = 'text',
            text = ~paste('Avg. Base Pay: $', Base.Pay)) %>%
  add_trace(x = ~Total.Pay, y = ~County.Name, name = "Total Pay", type = 'scatter',
            mode = "markers", marker = list(color = "purple"),
            hoverinfo = 'text',
            text = ~paste('Avg. Total Pay: $', Total.Pay)) %>%
  layout(
    title = "Benefits, Base Pay, and Total Pay in Bay Area Counties",
    xaxis = list(title = "Average Pay"),
    yaxis = list(title = FALSE),

    margin(list(l = 120))
  )

p

```

## Reasoning/Analysis
The notion behind this graph was to see if some counties paid their workers in different ways, as opposed to having a relatively equal distribution between benefits, base pay and total pay. San Francisco seemed to be the only unique county that we encountered. While their workers had proportionately far less Base Pay and Benefits, their total pay was still very competitive due to the large amount of overtime pay their workers logged. This would be an interesting avenue to pursue if we were to do future investigation into government jobs and *perhaps* government corruption

```{r 3rd Graph, echo=FALSE, warning=FALSE, Messege=FALSE}
#Bay Average
bay_av <- (num_son/tot_son + num_napa/tot_napa + num_solano/tot_solano + num_marin/tot_marin+ 
                        num_cc/tot_cc +num_sf/tot_sf +num_mateo/tot_mateo +num_cruz/tot_cruz+ 
                        num_clara/tot_clara + num_alam/tot_alam)/10 

#Physician Average
tot_phys <- dplyr::filter(merged.data, grepl('Physician', Job.Title))
tnphys <- nrow(tot_phys)
phys50 <- tot_phys$Base.Pay[tot_phys$Total.Pay...Benefits > 50000]
phys50 <- as.numeric(phys50)
phys50 <- matrix(data = phys50, ncol = 1)
phys_avg <- mean(phys50)
num_phys <- nrow(phys50)

#Nurse Average
tot_nurse <- dplyr::filter(merged.data, grepl('Nurse', Job.Title))
tnnurse  <- nrow(tot_nurse)
nurse50 <- tot_nurse$Base.Pay[tot_nurse$Total.Pay...Benefits > 50000]
nurse50 <- as.numeric(nurse50)
nurse50 <- matrix(data = nurse50, ncol = 1)
nurse_avg <- mean(nurse50)
num_nurse <- nrow(nurse50)

#Accountant
tot_acc <- dplyr::filter(merged.data, grepl('Account', Job.Title))
tnacc <- nrow(tot_acc)
acc50 <- tot_acc$Base.Pay[tot_acc$Total.Pay...Benefits > 50000]
acc50 <- as.numeric(acc50)
acc50 <- matrix(data = acc50, ncol = 1)
acc_avg <- mean(acc50)
num_acc <- nrow(acc50)

#attorney
tot_att <- dplyr::filter(merged.data, grepl('Attorney', Job.Title))
tnatt <- nrow(tot_att)
att50 <- tot_att$Base.Pay[tot_att$Total.Pay...Benefits > 50000]
att50 <- as.numeric(att50)
att50 <- matrix(data = att50, ncol = 1)
att_avg <- mean(att50)
num_att <- nrow(att50)

#janitor
tot_jan <- dplyr::filter(merged.data, grepl('Janitor', Job.Title))
tnjan <- nrow(tot_jan)
jan50 <- tot_jan$Base.Pay[tot_jan$Total.Pay...Benefits > 50000]
jan50 <- as.numeric(jan50)
jan50 <- matrix(data = jan50, ncol = 1)
jan_avg <- mean(jan50)
num_jan <- nrow(jan50)

#secretary
tot_sec <- dplyr::filter(merged.data, grepl('Secretary', Job.Title))
tnsec <- nrow(tot_sec)
sec50 <- tot_sec$Base.Pay[tot_sec$Total.Pay...Benefits > 50000]
sec50 <- as.numeric(sec50)
sec50 <- matrix(data = sec50, ncol = 1)
sec_avg <- mean(sec50)
num_sec <- nrow(sec50)

#engineer
tot_eng <- dplyr::filter(merged.data, grepl('Engineer', Job.Title))
tneng<- nrow(tot_eng)
eng50 <- tot_eng$Base.Pay[tot_eng$Total.Pay...Benefits > 50000]
eng50 <- as.numeric(eng50)
eng50 <- matrix(data = eng50, ncol = 1)
eng_avg <- mean(eng50)
num_eng <- nrow(eng50)

#Totals
emps <- c("Physician", "Nurse","Accountant","Attorney","Janitor","Secretary","Engineer")
emp_rats <- c(num_phys/tnphys,num_nurse/tnnurse,num_acc/tnacc,num_att/tnatt,num_jan/tnjan,num_sec/tnsec,num_eng/tneng)
all_emp <- data.frame(Occupations = c(emps), emp_rates = c(emp_rats), bay_rate = c(bay_av))

#Bar Graph

q <- plot_ly(all_emp, x = ~Occupations, y = ~emp_rates, type = 'bar', name = 'Occupation' ) %>%
  add_trace(y = ~bay_rate, name = 'Bay Area Average') %>%
  layout(  title = "Occupations Compared to Bay Area Average",
yaxis = list(range = c(.5, .95),title = 'Percentage Earning Over $50,000'), xaxis = list(title = FALSE), barmode = 'group')
q
```

## Reasoning/Analysis
For our final graph, we wanted to see how various occupations compared to the bay area average in terms of percentage of individuals per county being paid over #$50,000. Some of the results we saw were surprising, such as the occupation of attorney having above %93 making above 50K. Also, seeing that a larger percentage of janitors earning above this threshold when compared to nurses was also surprising. While nurses are caring for people and occasionally saving lives, they still do not on average make more than these janitors. 


